---
title: "R Notebook"
output:
  word_document: default
  pdf_document:
    latex_engine: xelatex
  html_notebook: default
always_allow_html: yes
---

```{r setup, include=FALSE}

# This chunk shows/hides the code in your final report. When echo = TRUE, the code
# is shown in the report. When echo = FALSE, the code is hidden from the final report.
# We would like to see your code, so please leave the setting as is during the course.
# This chunk will not show up in your reports, so you can safely ignore its existence.

knitr::opts_chunk$set(echo = TRUE)

```


The following is your first chunk to start with. Remember, you can add chunks using the menu
above (Insert -> R) or using the keyboard shortcut Ctrl+Alt+I. A good practice is to use
different code chunks to answer different questions. You can delete this comment if you like.

Other useful keyboard shortcuts include Alt- for the assignment operator, and Ctrl+Shift+M
for the pipe operator. You can delete these reminders if you don't want them in your report.


```{r}
# setwd("C:/Users/Avinav/UMD/Sem 2/BUDT758T/Data") #Don't forget to set your working directory before you start!

library("tidyverse")
library("tidymodels")
library("plotly")
library("skimr")
library("caret")
library("dplyr")
library("glmnet")
library("cowplot")
```

```{r}
# 1.
setwd("C:/Users/Avinav/UMD/Sem 2/BUDT758T/Data")

dfc <- read_csv("assignment3Carvana.csv")
skim(dfc)

```

```{r}

dfc

```

```{r}

#2.a
set.seed(52156)
dfcTrain <- dfc %>% sample_frac(0.65)
dfcTest <- dplyr::setdiff(dfc, dfcTrain)

```

```{r}

plot <- dfcTrain %>% 
  ggplot(aes(x = factor(BadBuy) , y = MMRAauction)) + geom_boxplot()
plot

```

```{r}

plot <- dfcTrain %>% 
  ggplot(aes(x = factor(BadBuy) , y = Age)) + geom_boxplot()
plot

```

```{r}

plot <- dfcTrain %>% 
  ggplot(aes(x = factor(BadBuy) , y = Odo)) + geom_boxplot()
plot

```

```{r}

# 2.b
dfcTrain %>% 
  group_by(Size) %>%
  summarise(GoodCars = length(BadBuy) - sum(BadBuy), Lemons = sum(BadBuy)) %>% 
  mutate(Percentage = Lemons/(GoodCars + Lemons)*100) %>% 
  arrange(desc(Percentage))

```

```{r}

# 3.a
fitLem <- lm(formula = BadBuy ~., data = dfcTrain)
summary(fitLem)

```

```{r}

resultsLem <- dfcTrain %>% mutate(Prediction = predict(fitLem, dfcTrain))
rmse(resultsLem, truth = BadBuy, estimate = Prediction)
```


```{r}
resultsLem <- dfcTest %>% mutate(Prediction = predict(fitLem, dfcTest))
rmse(resultsLem, truth = BadBuy, estimate = Prediction)

```

```{r}

# 3.c
resultsLem <-
	fitLem %>%
  predict(dfcTest, type = "response") %>% 
  bind_cols(dfcTest, predictedProb=.) %>% 
  select(BadBuy, predictedProb) %>% 
  mutate(predictedBadBuy = as.factor(ifelse(predictedProb > 0.5, 1, 0)), BadBuy = as.factor(BadBuy))
resultsLem %>% 
  conf_mat(truth = BadBuy, estimate = predictedBadBuy)

```

```{r}

# 3.d
resultsLem %>% 
  xtabs(~predictedBadBuy+BadBuy, .) %>% 
  confusionMatrix(positive = '1')

```

```{r}

dataGLM = data.frame(Auction="ADESA", Age=1, Make="HONDA",Color="SILVER", WheelType="Covers",Odo=10000, Size="LARGE",MMRAauction=8000, MMRAretail=10000)

predict(fitLem, dataGLM, type="response")

```


```{r}
# 4.a
dfcTrain <- dfcTrain %>% 
  mutate(BadBuy = as.factor(BadBuy))
dfcTest <- dfcTest %>% 
  mutate(BadBuy = as.factor(BadBuy))

resultsLemLR <-
	train(BadBuy ~ ., data = dfcTrain, family = "binomial", method = "glm") %>%
  predict(dfcTest, type = "raw") %>% 
  bind_cols(dfcTest, predictedClass=.)
skim(resultsLemLR)

```

```{r}

resultsLemLR %>% 
  group_by(Auction) %>% 
  tally()

```

```{r}

resultsLemLR %>% 
  group_by(Make) %>% 
  tally()

```

```{r}

resultsLemLR %>% 
  group_by(Color) %>% 
  tally()

```

```{r}

resultsLemLR %>% 
  group_by(WheelType) %>% 
  tally()

```

```{r}

resultsLemLR %>% 
  group_by(Size) %>% 
  tally()

```

```{r}

dfcNew <- dfc
dfcNew <- dfcNew %>% 
  mutate(Color = ifelse(Color == "NULL", "NOTAVAIL", Color)) 

dfMake <- dfcNew %>% 
  group_by(Make) %>% 
  tally() %>% 
  filter(n < 10)

dfcNew <- dfcNew %>%  
  mutate(Make = ifelse(Make %in% dfMake$Make, "OTHER", Make))

```

```{r}

dfcNew <- dfcNew %>% 
  mutate(BadBuy = as.factor(BadBuy))

set.seed(52156)
dfcTrain <- dfcNew %>% sample_frac(0.65)
dfcTest <- dplyr::setdiff(dfcNew, dfcTrain)

fitLemNew <- train(BadBuy ~ ., data = dfcTrain, family = "binomial", method = "glm") 
summary(fitLemNew)

```

```{r}

# 4.d
resultsLemNew <- fitLemNew %>% 
  predict(dfcTest, type = "raw") %>% 
  bind_cols(dfcTest, predictedClass=.)

resultsLemNew %>% 
  conf_mat(truth = BadBuy, estimate = predictedClass)

```

```{r}

resultsLemNew %>% 
  xtabs(~predictedClass+BadBuy, .) %>% 
  confusionMatrix(positive = '1') %>% 
  tidy()

```

```{r}

dataGLM = data.frame(Auction="ADESA", Age=1, Make="HONDA",Color="SILVER", WheelType="Covers",Odo=10000, Size="LARGE",MMRAauction=8000, MMRAretail=10000)

predict(fitLemNew, dataGLM, type="raw")

```


```{r}
#5.a
set.seed(123)
dfcTrain <- dfcNew %>% sample_frac(0.65)
dfcTest <- dplyr::setdiff(dfcNew, dfcTrain)

resultsLDA <- train(BadBuy ~ ., data = dfcTrain, family = "binomial", method = "lda", trControl=trainControl(method='cv', number=10)) %>%
  predict(dfcTest, type = "raw") %>% 
  bind_cols(dfcTest, predictedClass=.)
summary(resultsLDA)

```

```{r}

resultsLDA %>% 
  xtabs(~predictedClass+BadBuy, .) %>% 
  confusionMatrix(positive = '1')

```

```{r}

#5.b
set.seed(123)
fitKNN <- train(BadBuy~., data = dfcTrain, method = "knn", trControl=trainControl(method = "cv", number = 10), tuneLength = 20)

fitKNN

```

```{r}

fitKNN %>% 
  ggplot(aes(x=k, y=Accuracy))

```

```{r}

resultsKNN <- 
  fitKNN %>% 
  predict(dfcTest, type = 'raw') %>% 
  bind_cols(dfcTest, predictedClass=.)

resultsKNN %>% 
  xtabs(~predictedClass+BadBuy, .) %>% 
  confusionMatrix(positive='1')

```

```{r}

# 5.c
set.seed(123)
lambdaValues <- 10^seq(-5, 2, length = 100)
fitLasso <- train(BadBuy~., family='binomial', data=dfcTrain, method='glmnet', trControl=trainControl(method='cv', number=10), tuneGrid = expand.grid(alpha=1, lambda=lambdaValues))

summary(fitLasso)

```

```{r}

varImp(fitLasso)$importance %>% 
  rownames_to_column(var = "Variable") %>%
  mutate(Importance = scales::percent(Overall/100)) %>% 
  arrange(desc(Overall)) %>% 
  as_tibble()


```

```{r}
plot(varImp(fitLasso), top = 25)
```


```{r}

fitLasso$bestTune$lambda

```

```{r}

resultsLasso <- 
  fitLasso %>% 
  predict(dfcTest, type = 'raw') %>% 
  bind_cols(dfcTest, predictedClass=.)

resultsLasso %>% 
  xtabs(~predictedClass+BadBuy, .) %>% 
  confusionMatrix(positive='1')

```

```{r}

#5.d
set.seed(123)

resultsRidge <- 
  train(BadBuy~., family='binomial', data=dfcTrain, method='glmnet', trControl=trainControl(method='cv', number=10), tuneGrid = expand.grid(alpha=0, lambda=lambdaValues)) %>% 
  predict(dfcTest, type = 'raw') %>% 
  bind_cols(dfcTest, predictedClass=.)

resultsRidge %>% 
  xtabs(~predictedClass+BadBuy, .) %>% 
  confusionMatrix(positive='1')

```

```{r}

set.seed(123)

fitElasticNet <- train(BadBuy ~ ., family='binomial', data=dfcTrain, method='glmnet', trControl=trainControl(method='cv', number=10), tuneLength=10)

```

```{r}

varImp(fitElasticNet)$importance %>% 
  rownames_to_column(var = "Variable") %>%
  mutate(Importance = scales::percent(Overall/100)) %>% 
  arrange(desc(Overall)) %>% 
  as_tibble()

```

```{r}

plot(varImp(fitElasticNet), top = 25)

```

```{r}

fitElasticNet$bestTune

```

```{r}

plot(fitElasticNet$finalModel, xvar="lambda", label = TRUE)

```

```{r}

coef(fitElasticNet$finalModel, fitElasticNet$bestTune$lambda)

```

```{r}

resultsElasticNet <- 
  fitElasticNet %>%
  predict(dfcTest, type='raw') %>%
  bind_cols(dfcTest, predictedClass=.)

resultsElasticNet %>% 
  xtabs(~predictedClass+BadBuy, .) %>% 
  confusionMatrix(positive = '1')

```

```{r}

#5.e
set.seed(123)

resultsQDA <-
  train(BadBuy~., family = 'binomial', method = 'qda', data = dfcTrain, trControl = trainControl(method = "cv", number = 10)) %>% 
  predict(dfcTest, type = 'raw') %>% 
  bind_cols(dfcTest, predictedClass=.)

resultsQDA %>% 
  xtabs(~predictedClass+BadBuy, .) %>% 
  confusionMatrix(positive = '1')

```

```{r}

corData <- cor(dfcTrain[,c(2,6,8,9)])
corData

```


```{r}

resultsQDA <-
  train(BadBuy~.-Make-MMRAauction, family = 'binomial', method = 'qda', data = dfcTrain, trControl = trainControl(method = "cv", number = 10)) %>% 
  predict(dfcTest, type = 'raw') %>% 
  bind_cols(dfcTest, predictedClass=.)

resultsQDA %>% 
  xtabs(~predictedClass+BadBuy, .) %>% 
  confusionMatrix(positive = '1')

```

```{r}

options(yardstick.event_first = FALSE)

out1 <- 
  resultsLDA %>% 
  select(BadBuy, predictedClass) %>% 
  mutate(model = "m1")

out2 <- 
  resultsKNN %>%
  select(BadBuy, predictedClass) %>% 
  mutate(model = "m2")

out3 <- 
  resultsLasso %>%
  select(BadBuy, predictedClass) %>% 
  mutate(model = "m3")

out4 <- 
  resultsRidge %>%  
  select(BadBuy, predictedClass) %>% 
  mutate(model = "m4")

out5 <- 
  resultsElasticNet %>%
  select(BadBuy, predictedClass) %>% 
  mutate(model = "m5")

out6 <- 
  resultsQDA %>% 
  select(BadBuy, predictedClass) %>% 
  mutate(model = "m6")

outAll <- bind_rows(out1, out2, out3, out4, out5, out6)
```






