---
title: "R Notebook"
output:
  word_document: default
  pdf_document:
    latex_engine: xelatex
  html_notebook: default
always_allow_html: yes
---

```{r setup, include=FALSE}

# This chunk shows/hides the code in your final report. When echo = TRUE, the code
# is shown in the report. When echo = FALSE, the code is hidden from the final report.
# We would like to see your code, so please leave the setting as is during the course.
# This chunk will not show up in your reports, so you can safely ignore its existence.

knitr::opts_chunk$set(echo = TRUE)

```


The following is your first chunk to start with. Remember, you can add chunks using the menu
above (Insert -> R) or using the keyboard shortcut Ctrl+Alt+I. A good practice is to use
different code chunks to answer different questions. You can delete this comment if you like.

Other useful keyboard shortcuts include Alt- for the assignment operator, and Ctrl+Shift+M
for the pipe operator. You can delete these reminders if you don't want them in your report.


```{r}
# setwd("C:/Users/Avinav/UMD/Sem 2/BUDT758T/Data") #Don't forget to set your working directory before you start!

library("tidyverse")
library("tidymodels")
library("plotly")
library("skimr")
library("caret")
```

```{r}
setwd("C:/Users/Avinav/UMD/Sem 2/BUDT758T/Data")

dff <- read_csv("lab3FraminghamHeart.csv")
skim(dff)

```

```{r}

# i.
colsToFactor <- c('gender', 'education', 'currentSmoker', 'BPMeds', 'prevalentStroke', 'prevalentHyp', 'diabetes')

```
```{r}

# ii.
dff <- dff %>% 
  mutate_at(colsToFactor, ~factor(.))
str(dff)

```

```{r}

# 1.
plot <- dff %>% 
  ggplot(aes(x = TenYearCHD, y = sysBP, group = TenYearCHD)) + geom_boxplot()
#ggplotly(plot)
plot

```

```{r}

plot <- dff %>% 
  ggplot(aes(x = TenYearCHD, y = diaBP, group = TenYearCHD)) + geom_boxplot()
#ggplotly(plot)
plot

```

```{r}

plot <- dff %>% 
  ggplot(aes(x = TenYearCHD, y = totChol, group = TenYearCHD)) + geom_boxplot()
#ggplotly(plot)
plot

```

```{r}

# 2.i
set.seed(123)
dffTrain <- dff %>% sample_frac(0.7)
dffTest <- dplyr::setdiff(dff, dffTrain)

```

```{r}

# 2.ii
dffTrain %>% 
  group_by(gender) %>% 
  tally() %>% 
  mutate(pct = 100*n/sum(n))

```

```{r}

dffTest %>% 
  group_by(gender) %>% 
  tally() %>% 
  mutate(pct = 100*n/sum(n))

```


```{r}

dffTrain %>% 
  group_by(ageGroup = cut_interval(age, length = 10)) %>% 
  tally() %>% 
  mutate(pct = 100*n/sum(n))

```

```{r}

plot <- dffTrain %>% 
  ggplot(aes(x = age, fill = gender)) + geom_histogram(color = 'black')
ggplotly(plot)

```

```{r}

# Q3.
fitLPM <- lm(formula = TenYearCHD ~ ., data = dffTrain)
summary(fitLPM)

```

```{r}

car::vif(fitLPM)

```

```{r}

fitLPM <- lm(formula = TenYearCHD ~ . - currentSmoker, data = dffTrain)
summary(fitLPM)

```

```{r}

car::vif(fitLPM)

```

```{r}

# Q4
resultsLPM <-
	lm(formula = TenYearCHD ~ . - currentSmoker, data = dffTrain) %>%
  predict(dffTest, type = "response") %>% 
  bind_cols(dffTest, predictedProb=.) %>% 
  mutate(predictedClass = ifelse(predictedProb > 0.5, 1, 0))
skim(resultsLPM)

```

```{r}

dffTest %>% 
  group_by(TenYearCHD) %>% 
  tally()

```

```{r}

resultsLPM %>% 
  group_by(predictedClass) %>% 
  tally()

```

```{r}

dffTrain <- dffTrain %>% 
  mutate(TenYearCHD = factor(TenYearCHD))
dffTrain

```

```{r}

dffTest <- dffTest %>% 
  mutate(TenYearCHD = factor(TenYearCHD))
dffTest

```


```{r}

# Q5
fitLR <- glm(formula = TenYearCHD ~ . - currentSmoker, data = dffTrain, family = "binomial")
summary(fitLR)

```

```{r}

exp(coef(fitLR))

```

```{r}

# Q6
resultsLog <-
	glm(formula = TenYearCHD ~ . - currentSmoker, data = dffTrain, family = "binomial") %>%
  predict(dffTest, type = "response") %>% 
  bind_cols(dffTest, predictedProb=.) %>% 
  mutate(predictedClass = as.factor(ifelse(predictedProb > 0.5, 1, 0)))
skim(resultsLog)

```

```{r}

resultsLog %>% 
  group_by(predictedClass) %>% 
  tally()

```

```{r}

# Q7
resultsLog %>% 
  conf_mat(truth = as.numeric(as.factor(TenYearCHD)), estimate = predictedClass)

```

```{r}

# Q8
plot <- resultsLog %>% 
  ggplot(aes(x = age, y = predictedProb, color = gender, linetype = currentSmoker)) + 
  geom_point() + 
  geom_smooth() +
  labs(title = "Heart disease probability by age", x = "Age", y = "Predicted Probability")
plot

```

```{r}

plot <- resultsLog %>% 
  ggplot(aes(x = cigsPerDay, y = predictedProb, color = gender)) + 
  geom_point() + 
  geom_smooth() +
  labs(title = "Heart disease probability by cigarettes per day", x = "Cigarettes Per Day", y = "Predicted Probability")
plot

```

```{r}

plot <- resultsLog %>% 
  ggplot(aes(x = totChol, y = predictedProb, color = currentSmoker)) + 
  geom_point() + 
  geom_smooth() +
  labs(title = "Heart disease probability by total cholestrol level", x = "Total Cholestrol Level", y = "Predicted Probability")
plot

```

```{r}

plot <- resultsLog %>% 
  ggplot(aes(x = glucose, y = predictedProb, color = gender, linetype = currentSmoker)) + 
  geom_point() + 
  geom_smooth() +
  labs(title = "Heart disease probability by glucose level", x = "Glucose Level", y = "Predicted Probability")
plot

```

```{r}

# Q9
resultsLogCaret <-
	train(TenYearCHD ~ ., data = dffTrain, family = "binomial", method = "glm") %>%
  predict(dffTest, type = "raw") %>% 
  bind_cols(dffTest, predictedClass=.)
skim(resultsLogCaret)

```


```{r}

resultsLogCaret %>% 
  group_by(predictedClass) %>% 
  tally()

```

```{r}

resultsLogCaret %>% 
  xtabs(~predictedClass+TenYearCHD, .) %>% 
  confusionMatrix(positive = '1')

```

```{r}

# Q10
setwd("C:/Users/Avinav/UMD/Sem 2/BUDT758T/Data")

dfBP <- read_csv("lab3BancoPortugal.csv")
skim(dfBP)

```

```{r}

colsToFactor <- c('openedAccount', 'newcustomer', 'agegroup', 'job', 'marital', 'education', 'default', 'housing', 'loan', 'contact', 'month', 'day_of_week', 'poutcome')

dfBP <- dfBP %>% 
  mutate_at(colsToFactor, ~factor(.))
str(dfBP)

```

```{r}

set.seed(123)
dfBPTrain <- dfBP %>% sample_frac(0.7)
dfBPTest <- dplyr::setdiff(dfBP, dfBPTrain)

```

```{r}

resultsLogCaretBP <-
	train(openedAccount ~ ., data = dfBPTrain, family = "binomial", method = "glm") %>%
  predict(dfBPTest, type = "raw") %>% 
  bind_cols(dfBPTest, predictedClass=.)
skim(resultsLogCaretBP)

```

```{r}

resultsLogCaretBP %>% 
  xtabs(~predictedClass+openedAccount, .) %>% 
  confusionMatrix(positive = '1')

```

```{r}

car::vif(glm(formula = openedAccount ~ . - newcustomer, data = dfBPTrain, family = "binomial"))

```


```{r}

resultsLogCaretBP <-
	train(openedAccount ~ . - newcustomer - emp.var.rate - euribor3m, data = dfBPTrain, family = "binomial", method = "glm") %>%
  predict(dfBPTest, type = "raw") %>% 
  bind_cols(dfBPTest, predictedClass=.)

resultsLogCaretBP %>% 
  xtabs(~predictedClass+openedAccount, .) %>% 
  confusionMatrix(positive = '1')

```

```{r}

resultsLogCaretBP <-
	train(openedAccount ~ . - newcustomer - agegroup - marital - education - contact - poutcome - month, data = dfBPTrain, family = "binomial", method = "glm") %>%
  predict(dfBPTest, type = "raw") %>% 
  bind_cols(dfBPTest, predictedClass=.)

resultsLogCaretBP %>% 
  xtabs(~predictedClass+openedAccount, .) %>% 
  confusionMatrix(positive = '1')

```

```{r}


resultsLogCaretBP <-
	train(openedAccount ~ . - newcustomer - agegroup - marital - education, data = dfBPTrain, family = "binomial", method = "glm") %>%
  predict(dfBPTest, type = "raw") %>% 
  bind_cols(dfBPTest, predictedClass=.)

resultsLogCaretBP %>% 
  xtabs(~predictedClass+openedAccount, .) %>% 
  confusionMatrix(positive = '1')

```

