---
title: "R Notebook"
output:
  word_document: default
  pdf_document:
    latex_engine: xelatex
  html_notebook: default
always_allow_html: yes
---

```{r setup, include=FALSE}

# This chunk shows/hides the code in your final report. When echo = TRUE, the code
# is shown in the report. When echo = FALSE, the code is hidden from the final report.
# We would like to see your code, so please leave the setting as is during the course.
# This chunk will not show up in your reports, so you can safely ignore its existence.

knitr::opts_chunk$set(echo = TRUE)

```


The following is your first chunk to start with. Remember, you can add chunks using the menu
above (Insert -> R) or using the keyboard shortcut Ctrl+Alt+I. A good practice is to use
different code chunks to answer different questions. You can delete this comment if you like.

Other useful keyboard shortcuts include Alt- for the assignment operator, and Ctrl+Shift+M
for the pipe operator. You can delete these reminders if you don't want them in your report.


```{r}
#setwd("C:/") #Don't forget to set your working directory before you start!

library("tidyverse")
library("tidymodels")
library("plotly")
library("skimr")
library("lubridate")
library("car")
```

```{r}
setwd("C:/Users/Avinav/UMD/Sem 2/BUDT758T/Data")

dfw <-
  read_csv("walmartSales.csv")

```

```{r}

head(dfw)

```

```{r}
# Q1

fitCPI <- lm(formula = Weekly_Sales ~ CPI, data = dfw)
summary(fitCPI)

```

```{r}
# Q2

plot <- dfw %>% 
  filter(Store == 10) %>% 
  ggplot(aes(x = CPI, y = Weekly_Sales)) + geom_point() + geom_smooth(method = lm)
plot
#ggplotly(plot)


```

```{r}

plot <- dfw %>% 
  filter(Store == 11) %>% 
  ggplot(aes(x = CPI, y = Weekly_Sales)) + geom_point() + geom_smooth(method = lm)
plot
#ggplotly(plot)

```

```{r}

plot <- dfw %>% 
  filter(Store == 12) %>% 
  ggplot(aes(x = CPI, y = Weekly_Sales)) + geom_point() + geom_smooth(method = lm)
plot
#ggplotly(plot)

```

```{r}

plot <- dfw %>% 
  filter(Store == 13) %>% 
  ggplot(aes(x = CPI, y = Weekly_Sales)) + geom_point() + geom_smooth(method = lm)
plot
#ggplotly(plot)

```

```{r}
# Q3

plot <- dfw %>% 
  filter(year(Date) == 2012) %>% 
  ggplot(aes(x = CPI, y = Weekly_Sales)) + geom_point() + geom_smooth(method = lm)
plot


```

```{r}
# Q4

plot <- dfw %>% 
  filter( Store == 1, year(Date) == 2010) %>% 
  ggplot(aes(x = CPI, y = Weekly_Sales)) + geom_point() + geom_smooth(method = lm)
plot

```

```{r}
# Q5

fitCPISize <- lm(formula = Weekly_Sales ~ CPI + Size, data = dfw)
summary(fitCPISize)
anova(fitCPI, fitCPISize)
  
```

```{r}
# Q7

fitFull <- lm(formula = Weekly_Sales ~ CPI + Size + IsHoliday + Temperature + Fuel_Price + Unemployment, data = dfw)
summary(fitFull)
anova(fitCPISize, fitFull)

```

```{r}
# Q8

fitFullTemp <- lm(formula = Weekly_Sales ~ CPI + Size + IsHoliday + Temperature + I(Temperature^2) + Fuel_Price + Unemployment, data = dfw)
summary(fitFullTemp)

```

```{r}

dfw %>% ggplot(aes(x = Temperature, y = Weekly_Sales)) + 
  geom_smooth(method = lm, formula = y ~ x + I(x^2))

```

```{r}
# Q9

# a.
set.seed(333)

# b.
dfwTrain <- dfw %>% sample_frac(0.8)
dfwTest <- dplyr::setdiff(dfw, dfwTrain)
```

```{r}
# c.
fitOrg <- lm(formula = Weekly_Sales ~ CPI + Size + IsHoliday + Temperature + I(Temperature^2) + Fuel_Price + Unemployment, data = dfwTest)
summary(fitOrg)

```

```{r}
tidy(fitOrg)
```


```{r}
# d.
resultsOrg <- dfwTest %>% mutate(predictedSales = predict(fitOrg, dfwTest))
resultsOrg

```

```{r}
# e.
performance <- metric_set(rmse, mae)
performance(resultsOrg, truth = Weekly_Sales, estimate = predictedSales)

```

```{r}
# f.
fitOrgDate <- lm(formula = Weekly_Sales ~ CPI + Size + IsHoliday + Temperature + I(Temperature^2) + Fuel_Price + Unemployment + Date, data = dfwTest)
summary(fitOrgDate)

```

```{r}
resultsOrgDate <- dfwTest %>% mutate(predictedSales = predict(fitOrgDate, dfwTest))
performance(resultsOrgDate, truth = Weekly_Sales, estimate = predictedSales)

```

```{r}
anova(fitOrg, fitOrgDate)

```


```{r}
# g.
fitOrgNoUn <- lm(formula = Weekly_Sales ~ CPI + Size + IsHoliday + Temperature + I(Temperature^2) + Fuel_Price, data = dfwTest)
summary(fitOrgNoUn)

```

```{r}
resultsOrgNoUn <- dfwTest %>% mutate(predictedSales = predict(fitOrgNoUn, dfwTest))
performance(resultsOrgNoUn, truth = Weekly_Sales, estimate = predictedSales)

```

```{r}
anova(fitOrgNoUn, fitOrg)
```


```{r}
# Q10

set.seed(333)
dfwTrain <- dfw %>% sample_frac(0.8)
dfwTest <- setdiff(dfw, dfwTrain)
fitLog <- lm(formula = log(Weekly_Sales) ~ CPI + Size + IsHoliday + Temperature + I(Temperature^2) + Fuel_Price + Unemployment, data = dfwTrain)
summary(fitLog)

```

```{r}
resultsLog <- dfwTest %>% mutate(predictedSales = predict(fitLog, dfwTest))
performance(resultsLog, truth = Weekly_Sales, estimate = exp(predictedSales))

```

```{r}
# b.
plot(fitOrg)

```

```{r}
plot(fitLog)

```

```{r}
# Bonus question

set.seed(333)
dfwTrain <- dfw %>% sample_frac(0.8)
dfwTest <- setdiff(dfw, dfwTrain)
fitSalesSqFoot <- lm(formula = Weekly_Sales/Size ~ CPI + Size + IsHoliday + Temperature + I(Temperature^2) + Fuel_Price + Unemployment, data = dfwTest)
summary(fitSalesSqFoot)

```

```{r}
resultsSalesSqFoot <- dfwTest %>% mutate(predictedSales = predict(fitSalesSqFoot, dfwTest))
performance(resultsSalesSqFoot, truth = Weekly_Sales, estimate = predictedSales*Size)

```

```{r}
set.seed(333)
dfwTrain <- dfw %>% sample_frac(0.8)
dfwTest <- setdiff(dfw, dfwTrain)
fitSalesSqFoot <- lm(formula = Weekly_Sales/Size ~ CPI + IsHoliday + Temperature + I(Temperature^2) + Fuel_Price + Unemployment, data = dfwTest)
summary(fitSalesSqFoot)

```

```{r}
resultsSalesSqFoot <- dfwTest %>% mutate(predictedSales = predict(fitSalesSqFoot, dfwTest))
performance(resultsSalesSqFoot, truth = Weekly_Sales, estimate = predictedSales*Size)

```

