---
title: "R Notebook"
output:
  word_document: default
  pdf_document:
    latex_engine: xelatex
  html_notebook: default
always_allow_html: yes
---

```{r setup, include=FALSE}

# This chunk shows/hides the code in your final report. When echo = TRUE, the code
# is shown in the report. When echo = FALSE, the code is hidden from the final report.
# We would like to see your code, so please leave the setting as is during the course.
# This chunk will not show up in your reports, so you can safely ignore its existence.

knitr::opts_chunk$set(echo = TRUE)

```


The following is your first chunk to start with. Remember, you can add chunks using the menu
above (Insert -> R) or using the keyboard shortcut Ctrl+Alt+I. A good practice is to use
different code chunks to answer different questions. You can delete this comment if you like.

Other useful keyboard shortcuts include Alt- for the assignment operator, and Ctrl+Shift+M
for the pipe operator. You can delete these reminders if you don't want them in your report.


```{r}
# setwd("C:/") #Don't forget to set your working directory before you start!

library("tidyverse")
library("tidymodels")
library("plotly")
library("skimr")
library("modelr")
library("dplyr")
library("lubridate")
```

```{r}

setwd("C:/Users/Avinav/UMD/Sem 2/BUDT758T/Data")

dfbOrg <-
  read_csv("assignment2BikeShare.csv")

```

```{r}
skim(dfbOrg)

```

```{r}

# Q1.a.i
dfbOrg <- dfbOrg %>% 
  mutate(COUNT = CASUAL + REGISTERED)

dfbOrg

```

```{r}

# Q1.a.ii
dfbOrg <- dfbOrg %>% 
  mutate(MONTH = months(DATE))

dfbOrg

```

```{r}

# Q1.b
dfbStd <- dfbOrg
dfbStd$TEMP <- scale(dfbOrg$TEMP)
dfbStd$ATEMP <- scale(dfbOrg$ATEMP)
dfbStd$HUMIDITY <- scale(dfbOrg$HUMIDITY)
dfbStd$WINDSPEED <- scale(dfbOrg$WINDSPEED)

head(dfbStd)

```

```{r}

# Q2
fitAll <- lm(formula = COUNT ~. - COUNT, data = dfbStd)
summary(fitAll)

```

```{r}

# Q3.a
dfbOrg <- dfbOrg %>% 
  mutate(BADWEATHER = ifelse(WEATHERSIT == 3 | WEATHERSIT == 4, "YES", "NO"))

dfbOrg

```

```{r}

# Q3.b
plot <- dfbOrg %>% 
  ggplot(aes(x = ATEMP, y = COUNT)) + geom_point(aes(color = BADWEATHER))
plot

```

```{r}

# Q3.c
plot <- dfbOrg %>% 
  ggplot(aes(x = ATEMP, y = CASUAL)) + geom_point(aes(color = BADWEATHER))
plot

```

```{r}

plot <- dfbOrg %>% 
  ggplot(aes(x = ATEMP, y = REGISTERED)) + geom_point(aes(color = BADWEATHER))
plot

```

```{r}

# Q3.c.iv
plot <- dfbOrg %>% 
  ggplot(aes(x = ATEMP, y = COUNT)) + geom_point() + geom_smooth()
plot

```

```{r}

# Q4
fitOrg <- lm(formula = COUNT ~ MONTH + WEEKDAY + BADWEATHER + TEMP + ATEMP + HUMIDITY, data = dfbOrg)
summary(fitOrg)

```

```{r}

# Q5
plot(fitOrg)

```

```{r}

dfbOrg %>% 
  add_residuals(fitOrg) %>% 
  ggplot(aes(DATE, resid)) + geom_line()

```

```{r}

car::vif(fitOrg)

```

```{r}

fitOrg <- lm(formula = COUNT ~ MONTH + WEEKDAY + BADWEATHER + TEMP + HUMIDITY, data = dfbOrg)
plot(fitOrg)

```

```{r}

dfbOrg %>% 
  add_residuals(fitOrg) %>% 
  ggplot(aes(DATE, resid)) + geom_line()

```


```{r}

car::vif(fitOrg)

```

```{r}

# Q6
fit <- lm(formula = COUNT ~ BADWEATHER, data = dfbOrg)
summary(fit)

```

```{r}

# Q6.c
fit <- lm(formula = COUNT ~ BADWEATHER * WEEKDAY, data = dfbOrg)
summary(fit)

```

```{r}

# Q7
# a.
set.seed(333)

# b.
dfbTrain <- dfbOrg %>% sample_frac(0.8)
dfbTest <- dplyr::setdiff(dfbOrg, dfbTrain)

```

```{r}

# c.i
fitOrg <- lm(formula = COUNT ~ MONTH + WEEKDAY + BADWEATHER + TEMP + HUMIDITY, data = dfbTrain)
summary(fitOrg)

```

```{r}

detach('package:modelr', unload=TRUE)
resultsOrg <- dfbTest %>% mutate(PREDICTEDCOUNT = predict(fitOrg, dfbTest))
performance <- metric_set(rmse, mae)
performance(resultsOrg, truth = COUNT, estimate = PREDICTEDCOUNT)

```

```{r}

fitNew <- lm(formula = COUNT ~ MONTH + WEEKDAY + BADWEATHER + TEMP + HUMIDITY + WINDSPEED, data = dfbTrain)
summary(fitNew)

```

```{r}

resultsNew <- dfbTest %>% mutate(PREDICTEDCOUNT = predict(fitNew, dfbTest))
performance(resultsNew, truth = COUNT, estimate = PREDICTEDCOUNT)

```

```{r}

anova(fitOrg, fitNew)

```

```{r}

# Q8
dfbTrain <- dfbOrg %>% 
  filter(year(DATE) == 2011)
dfbTest <- dfbOrg %>% 
  filter(year(DATE) == 2012)
fitTime <- lm(formula = COUNT ~ MONTH + WEEKDAY + BADWEATHER + TEMP + HUMIDITY, data = dfbTrain)
summary(fitTime)

```

```{r}

resultsTime <- dfbTest %>% mutate(PREDICTEDCOUNT = predict(fitTime, dfbTest))
performance(resultsTime, truth = COUNT, estimate = PREDICTEDCOUNT)

```

```{r}

dfbTrain <- dfbOrg %>% 
  filter(year(DATE) == 2011 | (year(DATE) == 2012 & month(DATE) <= 6))
dfbTest <- dfbOrg %>% 
  filter(year(DATE) == 2012 & month(DATE) > 6)
fitTimeNew <- lm(formula = COUNT ~ MONTH + WEEKDAY + BADWEATHER + TEMP + HUMIDITY, data = dfbTrain)
summary(fitTimeNew)

```

```{r}

resultsTimeNew <- dfbTest %>% mutate(PREDICTEDCOUNT = predict(fitTimeNew, dfbTest))
performance(resultsTimeNew, truth = COUNT, estimate = PREDICTEDCOUNT)

```


